Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-23 ICML-submission/appendix-experiment.tex

\section{Supplementary Materials for Section~\ref{sec:experiments}}
\label{sec:supp-to-experiment}

However, in 2D, the samples in Figure~\ref{fig:strong-points} do not satisfy the strong linear separability defined in Definition~\ref{definition:weak-linear-separability}, which requires that any class can be separated from all the others by a hyperplane that \textit{passes through the origin}. We therefore raise the dimensionality to 3D by adding a constant feature to all samples. This is equivalent to allowing the separating hyperplanes to have a bias, so that they do not need to pass through the origin. Thus Figure~\ref{fig:strong-points} becomes strongly linearly separable. To better understand how each algorithm works, we plot in Figure~\ref{fig:banditron-points},~\ref{fig:linearova-points},~\ref{fig:rationalova-points} the final decision boundary that is learned by each algorithm.

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_banditron_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_banditron_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{\textsc{Banditron}'s decision boundary}
     \label{fig:banditron-points}
    % \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$. }
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_linear_ova_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_linear_ova_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Our Algorithm with linear kernel (Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})'s decision boundary}
    \label{fig:linearova-points}
   % \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$. }
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_rational_ova_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_rational_ova_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Our Algorithm with rational kernel (Algorithm~\ref{algorithm:kernelized})'s decision boundary}
    \label{fig:rationalova-points}
    % \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$. }
\end{figure}

>>>>

\section{Supplementary Materials for Section~\ref{section:experiments}}
\label{section:supp-to-experiment}

To better understand how each algorithm works, we plot in
Figure~\ref{fig:banditron-points},~\ref{fig:linearova-points},~\ref{fig:rationalova-points}
the final decision boundary that is learned by each algorithm.

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_banditron_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_banditron_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{\textsc{Banditron}'s decision boundary}
     \label{fig:banditron-points}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_linear_ova_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_linear_ova_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Our Algorithm with linear kernel (Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})'s decision boundary}
    \label{fig:linearova-points}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_rational_ova_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_rational_ova_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Our Algorithm with rational kernel (Algorithm~\ref{algorithm:kernelized})'s decision boundary}
    \label{fig:rationalova-points}
\end{figure}

### 2019-01-23 ICML-submission/introduction.tex

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a repeated game between a learner and an adversary. At each timestep
$t$, the adversary chooses a labeled example $(x_t, y_t)$ and reveals $x_t\in
\R^d$ to the learner. Upon receiving $x_t$, the learner makes a prediction
$\widehat{y}_t$ and receives feedback. In contrast with the standard
full-information setting, where the feedback given is the correct label $y_t$,
here the feedback is only a binary indicator of whether the prediction was
correct or not. The protocol of the problem is formally stated below.

@@

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect: first, their Theorem 2 implies a mistake upper bound of
$\frac{R}{\gamma}$ with sufficiently large $C$, which contradicts our lower
bound in Theorem~\ref{theorem:strongly-separable-examples-mistake-lower-bound};
second, their Lemma 4 cannot be directly applied to the bandit setting, as the
algorithm only makes updates on a subset of the $w_i$'s.} We design a simple and
efficient algorithm
(Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})
that makes at most $O(K (R/\gamma)^2)$ mistakes in expectation. Its memory
complexity and per-round time complexity are both $O(dK)$. The algorithm can be
viewed as running $K$ copies of the \textsc{Binary Perceptron} algorithm, one
copy for each class. We prove that any (possibly randomized) algorithm must make
$\Omega(K (R/\gamma)^2)$ mistakes in the worst case. The extra $O(K)$
multiplicative factor in the mistake bound, as compared to the full-information
setting, is the price we pay for the bandit feedback, or more precisely, the
lack of full-information feedback.

>>>>

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a repeated game between a learner and an adversary. At each time step
$t$, the adversary chooses a labeled example $(x_t, y_t)$ and reveals $x_t\in
\R^d$ to the learner. Upon receiving $x_t$, the learner makes a prediction
$\widehat{y}_t$ and receives feedback. In contrast with the standard
full-information setting, where the feedback given is the correct label $y_t$,
here the feedback is only a binary indicator of whether the prediction was
correct or not. The protocol of the problem is formally stated below.

@@

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their
Theorem 2 is incorrect: first, their Theorem 2 implies a mistake upper bound of
$\frac{R}{\gamma}$ with sufficiently large $C$, which contradicts our lower
bound in Theorem~\ref{theorem:strongly-separable-examples-mistake-lower-bound};
second, their Lemma 4 cannot be directly applied to the bandit setting, as the
algorithm only makes updates on a subset of the $w_i$'s.} We design a simple and
efficient algorithm
(Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})
that makes at most $O(K (R/\gamma)^2)$ mistakes in expectation. Its memory
complexity and per-round time complexity are both $O(dK)$. The algorithm can be
viewed as running $K$ copies of the \textsc{Binary Perceptron} algorithm, one
copy for each class. We prove that any (possibly randomized) algorithm must make
$\Omega(K (R/\gamma)^2)$ mistakes in the worst case. The extra $O(K)$
multiplicative factor in the mistake bound, as compared to the full-information
setting, is the price we pay for the bandit feedback, or more precisely, the
lack of full-information feedback.

### 2019-01-23 ICML-submission/main-all.tex
Switch to one .tex file
Move proofs for strongly separable data to appendix
a version within page limit
Add NP-hardness of weak labeling problem
nn alg
Update main-all.tex
Moved the proof of Theorem 6 (originally Appendix C, now Appendix E) after the proof of Lemma 10 and the proof of Theorem 8 and 9 (previously Appendices D and E, now Appendices C and D)
add experiments

\input{title-and-abstract}
\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}
\input{algorithm-for-strongly-linearly-separable-data}
\input{from-linear-separability-to-strong-linear-separability}

@@

\input{appendix-multiclass-perceptron}
\input{appendix-proofs-of-polynomial-approximation-theorems}

>>>>

\input{title-and-abstract}
\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}
\input{algorithm-for-strongly-linearly-separable-data}
\input{from-weak-separability-to-strong-separability}
\input{empirical-comparison}

@@

\input{appendix-multiclass-perceptron}
\input{appendix-proofs-for-strongly-separable-examples}
\input{appendix-poly-norm-to-rkhs-norm}
\input{appendix-proofs-of-polynomial-approximation-theorems}
\input{appendix-margin-transformation}
\input{appendix-experiment}
\input{appendix-nearest-neighbor-algorithm}
\input{np-hardness-of-the-weak-labeling-problem}
\input{mistake-lower-bound-for-ignorant-algorithms}
